Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-31 tests-sodshocktube.tex
sweep 1: typo changed

We begin the paper test suite with the classic Sod shock tube problem
\citep{Sod78}, which provides a good test of a hydrodynamical solver's
ability to resolve a clean Riemann problem with clear separation
between the three waves.  These waves include a rarefaction fan, a
contact discontinuity, and a moderate shock.  The initial state is 
$\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on the left of the boundary at 
$x=0.5$ and $\rho_{\rm R}, P_{\rm R} = 0.125, 0.1$ on the right.
All velocities are initially zero.  In
Figure~\ref{fig.sodshocktube}, we show the density solution at the
final time ($t=0.25$) for three of our hydro solvers -- the spatially third-order
PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
100 cells across the domain, (which is a relatively standard choice in
code method papers) and show solutions both with and without adaptive
mesh refinement.  Without AMR (top row in Figure \ref{fig.sodshocktube}), it is
clear that the PPM scheme produces by far the cleanest solution with
all wave families crisply reproduced (in particular, the contact
discontinuity and shock).  \zeus\ and MUSCL produce similar results,
with MUSCL doing a slightly better job on the rarefaction fan.  In Figure
\ref{fig.sodshocktube}, we also show the integrated absolute deviation from 
the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-Exact(x_i)|$ 
, as well as the maximum deviation,
$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-Exact(x_i)| \right\}$.  
These numbers confirm the qualitative differences noted
previously.

>>>>

We begin the paper test suite with the classic one-dimensional Sod shock tube problem
\citep{Sod78}, which provides a good test of a hydrodynamical solver's
ability to resolve a clean Riemann problem with clear separation
between the three resultant waves.  These waves consist of a rarefaction fan, a
contact discontinuity, and a moderate shock.  The initial state is 
$\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on the left of the boundary at 
$x=0.5$ and $\rho_{\rm R}, P_{\rm R} = 0.125, 0.1$ on the right.
All velocities are initially zero.  In
Figure~\ref{fig.sodshocktube}, we show the density solution at the
final time ($t=0.25$) for three of our hydro solvers -- the spatially third-order
PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
100 cells across the domain, (which is a relatively standard choice in
code method papers) and show solutions both with and without adaptive
mesh refinement.  Without AMR (top row in Figure \ref{fig.sodshocktube}), it is
clear that the PPM scheme produces by far the cleanest solution with
all wave families crisply reproduced (in particular, the contact
discontinuity and shock).  \zeus\ and MUSCL produce similar results,
with MUSCL doing a slightly better job on the rarefaction fan.  In Figure
\ref{fig.sodshocktube}, we also show the integrated absolute deviation from 
the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-{\rm Exact}(x_i)|$ 
, as well as the maximum deviation,
$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-{\rm Exact}(x_i)| \right\}$.  
These numbers confirm the qualitative differences noted
previously.

### 2013-05-31 tests-vandv.tex
sweep 1: typo changed

\enzo\ is capable of simulating a large variety of problems types,
with all but a few of these types requiring only a parameter file as an
input.  The notable exception is the cosmology simulation, which takes
as input initial conditions created by other codes.  The suite of test
problems spans a wide range in complexity.  At one end of this spectrum
are simple problems that utilize only a single component of \enzo\ and
for which analytic solutions exist for comparison with the simulation
results.  At the opposite end are problems that exercise a large
portion of \enzo's machinery in concert.  Together, the available
problem types fully cover all of \enzo's functionality.  This enables
them to serve as a vehicle for verifying that the code's behavior
remains stable over time and after modifcation of the codebase.

@@

After the test problems are selected by the test runner script, they
are run in succession using either the \enzo\ executable contained
within that distribution or an external executable built from another \enzo\
version and specified by the user.  This allows for any version of the
code to be tested with an identical set of test problems.  After
running the test problem simulations, the test runner then performs a
series of basic analysis tasks using the \texttt{yt} analysis toolkit
\citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  The default
analysis performed on all test problems includes calculation of
various statistics, such as extrema; mean; and variance, on the fields
present in the output data.  Custom analysis provided by scripts that
accompany the test problem parameter files is run for special cases,
such as when an analytical solution exists that can be compared
against the simulation data.  After the analysis is performed, the
results are compared against a set of gold standard results that are
maintained on a website and downloaded on the fly by the test runner
script.  Alternately, results from any version of \enzo\ can be stored
locally and compared to any other version of the code.  In
Section~\ref{sec.tests.problems}, we describe some of the key test
problems that are used to verify proper behavior.  All of these test
problems, as well as the scripts to generate the figures from them,
are available in the \enzo\ test suite, allowing the interested reader
to regenerate all of the figures in Section~\ref{sec.tests.compare}
using the \textit{'methodpaper'} suite of test problems.

@@

\enzo\ has been involved in multiple cosmological code comparisons,
including the Santa Barbara Cluster code comparison project
\citep{SantaBarbara}, a large comparison of N-body simulations
\citep{2008CS&D....1a5003H}, as well as several direct comparisons
between \enzo\ and the GADGET SPH code under a variety of
circumstances, including N-body and adiabatic hydrodynamics
\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
The \enzo\ code typically has a more difficult time resolving
small-scale self-gravitating structures (for an equivalent dark matter
particle count and nominally equivalent force resolution), but does
comparably well as a tree-based code for larger structure, and is
typically superior in terms of resolving fluid features due to its
higher-order (and artificial viscosity-free) Piecewise Parabolic
Method hydrodynamics solver.  When examining classical test cases such
as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
clusters with very similar density, temperature, and entropy profiles
to other grid-based codes that use Godunov-type hydro methods, which
systematically differed from particle-based codes using SPH in this
code comparison.  Similarly, in galaxy cluster simulations that look
specifically at the properties of cosmological shocks
\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
similar to other high-order grid-based hydrodynamics codes, and far
superior performance of \enzo\ is observed (in terms of resolution of
fluid features and shock detection) in low-density regions when
compared to a particle-based code.  In tests of the Lyman-alpha forest
that include radiative cooling and a uniform metagalactic ultraviolet
background, \enzo\ and Gadget provide results on metrics such as the
matter power spectrum that are comparable to within 5\%
\citep{2007MNRAS.374..196R}.

@@

Three other comparisons between the \enzo\ code and other simulation
tools have been performed that focus on physics other than gravity and
fluid flow.  The flux-limited diffusion radiation transport scheme was
measured against several test problems by \citet{IlievEtAl2009},
which involved tests with and without analytical solutions.  \enzo\ produced
results similar to both the analytical solutions and results obtained by 
other codes.  We note, however, that there were minor differences throughout
the comparison between various codes, and the majority of the codes differred
in at least a subset of the tests.
\citet{2011ApJ...726...55T} shows the results of varied reaction rate
coefficients for the formation of molecular hydrogen via the 3-body
process in both the \enzo\ and Gadget-2 codes, observing similar trends between
the two codes. However, at nominally equivalend resolution
where the particle and cell gas masses are comparable, 
\enzo\ simulations typically displayed a
substantially higher level of gas structure.  This is unsurprising due
to \enzo's higher-order hydro solver.  Finally,
\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
the context of common-envelope binary stellar interactions.  The
authors show that the codes display reasonable convergence properties
as a function of simulation resolution, and also agree quite well with
each other. However, the observed mass-loss rates do not agree
particularly well with observations.

>>>>

\enzo\ is capable of simulating a large variety of problems types,
with all but a few of these types requiring only a parameter file as an
input.  The notable exception is the cosmology simulation, which takes
as input initial conditions created by other codes.  The suite of test
problems spans a wide range in complexity.  At one end of this spectrum
are simple problems that utilize only a single component of \enzo\ and
for which analytic solutions exist for comparison with the simulation
results.  At the opposite end are problems that exercise a large
portion of \enzo's machinery in concert.  Together, the available
problem types fully cover all of \enzo's functionality.  This enables
them to serve as a vehicle for verifying that the code's behavior
remains stable over time and after modification of the codebase.

@@

After the test problems are selected by the test runner script, they
are run in succession using either the \enzo\ executable contained
within that distribution or an external executable built from another \enzo\
version and specified by the user.  This allows for any version of the
code to be tested with an identical set of test problems.  After
running the test problem simulations, the test runner then performs a
series of basic analysis tasks using the \texttt{yt} analysis toolkit
\citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  The default
analysis performed on all test problems includes calculation of
various statistics, such as extrema, mean and variance, on the fields
present in the output data.  Custom analysis provided by scripts that
accompany the test problem parameter files is run for special cases,
such as when an analytical solution exists that can be compared
against the simulation data.  After the analysis is performed, the
results are compared against a set of gold standard results that are
maintained on a website and downloaded on the fly by the test runner
script.  Alternately, results from any version of \enzo\ can be stored
locally and compared to any other version of the code.  In
Section~\ref{sec.tests.problems}, we describe some of the key test
problems that are used to verify proper behavior.  All of these test
problems, as well as the scripts to generate the figures from them,
are available in the \enzo\ test suite, allowing the interested reader
to regenerate all of the figures in Section~\ref{sec.tests.compare}
using the \textit{`methodpaper'} suite of test problems.

@@

\enzo\ has been involved in multiple cosmological code comparisons,
including the Santa Barbara Cluster code comparison project
\citep{SantaBarbara}, a large comparison of N-body simulations
\citep{2008CS&D....1a5003H}, as well as several direct comparisons
between \enzo\ and the GADGET SPH code in a variety of
set-ups, including N-body and adiabatic hydrodynamics
\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
Compared to the other codes involved in these projects, the \enzo\ code typically has a more difficult time resolving
small-scale self-gravitating structures (for an equivalent dark matter
particle count and nominally equivalent force resolution), but does
comparably well as a tree-based code for larger structure, and is
typically superior in terms of resolving fluid features due to its
higher-order (and artificial viscosity-free) Piecewise Parabolic
Method hydrodynamics solver.  When examining classical test cases such
as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
clusters with very similar density, temperature, and entropy profiles
to other grid-based codes that use Godunov-type hydro methods, which
systematically differed from particle-based codes using SPH in this
code comparison.  Similarly, in galaxy cluster simulations that look
specifically at the properties of cosmological shocks
\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
similar to other high-order grid-based hydrodynamics codes, and a far
superior performance of \enzo\ is observed (in terms of resolution of
fluid features and shock detection) in low-density regions when
compared to a particle-based code.  In tests of the Lyman-alpha forest
that include radiative cooling and a uniform metagalactic ultraviolet
background, \enzo\ and Gadget provide results on metrics such as the
matter power spectrum that are comparable to within 5\%
\citep{2007MNRAS.374..196R}.

@@

Three other comparisons between the \enzo\ code and other simulation
tools have been performed that focus on physics other than gravity and
fluid flow.  The flux-limited diffusion radiation transport scheme was
measured against several test problems by \citet{IlievEtAl2009},
which involved tests with and without analytical solutions.  \enzo\ produced
results similar to both the analytical solutions and results obtained by 
other codes.  We note, however, that there were minor differences throughout
the comparison between various codes, and the majority of the codes differed
in at least a subset of the tests.
\citet{2011ApJ...726...55T} shows the results of varied reaction rate
coefficients for the formation of molecular hydrogen via the 3-body
process in both the \enzo\ and Gadget-2 codes, observing similar trends between
the two codes. However, at nominally equivalent resolution
where the particle and cell gas masses are comparable, 
\enzo\ simulations typically displayed a
substantially higher level of gas structure.  This is unsurprising due
to \enzo's higher-order hydro solver.  Finally,
\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
the context of common-envelope binary stellar interactions.  The
authors show that the codes display reasonable convergence properties
as a function of simulation resolution, and also agree quite well with
each other. However, the observed mass-loss rates do not agree
particularly well with observations.
